---
title: "Prediction of beer consumption in Sao Paulo Brazil"
output:
  html_document
    
---

<h5><b>Loading dependent libraries:</b></h5>
```{r echo=TRUE, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(readr)
library(caret)
library(stringr)
library(purrr)
library(lubridate)
library(corrplot)
library(caretEnsemble)
```

<h4><b>Here we import the data using the readr library.</b></h4>
<p>In our raw data commas are used as decimal points. readr reads the temperature data as integers and  ignores the commas so we import the numeric values as characters and will  convert them later on after replacing the commas with a decimal point</p>
```{r}
beerdata<-read_csv('Consumo_cerveja.csv',col_types = cols(
  
  `Temperatura Media (C)` = col_character(),
  `Temperatura Minima (C)` = col_character(),
  `Temperatura Maxima (C)` = col_character(),
  `Precipitacao (mm)` = col_character(),
  `Final de Semana` = col_character()
))
```




```{r}
str(beerdata)
```
<h5><b>Converting the column names  from Spanish to English</b> </h5>
```{r}
colnames(beerdata)<- c('date','mean_temp','min_temp','max_temp','rainfall','endofweek','beerconsumption')
```
```{r}
head(beerdata)
```

<h5><b>Trimming the dataset to include only the first 365 rows
as the data only pertains to one year and the remainig values are NULL.</b></h5>
```{r}
beerdata<-beerdata[ 1:365,]
tail(beerdata)
```


<h5><b>Replacing commas with decimal points</b></h5>
```{r}
removecomma <- function(x){ 
x<-str_replace(x,',','.') 
return(x) 
}

```

```{r}
beerdata$min_temp<-beerdata$min_temp%>%map_chr(removecomma)
beerdata$max_temp<-beerdata$max_temp%>%map_chr(removecomma)
beerdata$mean_temp<-beerdata$mean_temp%>%map_chr(removecomma)
beerdata$rainfall<-beerdata$rainfall%>%map_chr(removecomma)

```

```{r}
for(i in 2:5){
  beerdata[[i]]<-as.double(beerdata[[i]])
}
```
```{r}
head(beerdata)
```

<h5><b>Checking for NULL values</b></h5>

```{r}
sum(is.na(beerdata))
```


<h5> <b>It seems there is higher beer consumtion on weekends</b></h5>
```{r}
ggplot(data=beerdata,aes(x=endofweek,y=beerconsumption))+geom_boxplot(aes(fill=endofweek))+

scale_fill_brewer(type='div',palette = 4)+theme_minimal()
  
```

<h5><b>Adding day of week as a new column</b></h5>
```{r}
beerdata<-beerdata%>%mutate(dayofweek=wday(beerdata$date,label = TRUE))
```

<h5><b>There is a marked increase in beer consumtion on SATUDRAY
and SUNDAY</b></h5> 
```{r}
ggplot(data=beerdata,aes(x=dayofweek,y=beerconsumption))+geom_boxplot(aes(fill=dayofweek))+

scale_fill_brewer(palette = 'Spectral')+theme_bw()
```
<h5> <b>There seems to be a positive relationship between temperature and beer consumption fo evey day of the week.</b></h5>
```{r}
ggplot(data=beerdata,aes(x=mean_temp,y=beerconsumption))+geom_point(aes(color=mean_temp))+
  geom_smooth()+facet_wrap(~dayofweek)+scale_color_gradient(low="yellow", high="red")



```
<h5> <b>Since most days of the year in So Paulo have little o no rainfall,it is hard to infer any relationship between rainfall and beer consumption </b></h5> 
```{r}
ggplot(data=beerdata,aes(x=rainfall,y=beerconsumption))+geom_point(aes(color=rainfall))+
scale_color_gradient(low="#00a3cc", high="#002d39")
```

<h5><b>Adding the month to the data we notice that the consumption of beer dips towards the middle of the year but not by much</b></h5>
```{r}
beerdata%>%mutate(monthofyear=month(date,label = TRUE))%>%group_by(monthofyear)%>%summarise(avg_beer_consumption=mean(beerconsumption))%>%ggplot(aes(x=monthofyear,y=avg_beer_consumption,fill=avg_beer_consumption))+geom_bar(stat = 'identity')
```

<h5><b> None of the features are too correlated with beer consumtion</b></h5>
```{r}
correlation<-cor(beerdata%>%select(-c(date,endofweek,dayofweek)))
corrplot(correlation, method="number")

```


<h4><b>Finding important features using a random forest model</b></h4>
```{r}
control <- trainControl(method="repeatedcv", number=5, repeats=3)

model <- train(beerconsumption~., data=beerdata%>%select(-c(date,dayofweek)), method="rf", preProcess="scale", trControl=control)
importance <- varImp(model, scale=FALSE)

print(importance)

plot(importance)
```

<h5><b>Choosing the temperature variables and end-of-week as the final features and split the data into test and train datasets</b></h5>
```{r}
beerdata<-beerdata%>%select(max_temp,endofweek,min_temp,mean_temp,beerconsumption)

trainIndex <- createDataPartition(beerdata$beerconsumption, p = .8, 
                                  list = FALSE, 
                                  times = 1)

Train <- beerdata[ trainIndex,]
Test  <- beerdata[-trainIndex,]

head(Test)
```

<h5><b>Running 5 fold cross validation on five different models</b></h5>
<ul>
<li>linear regression</li>
<li>LASSO</li>
<li>RIDGE</li>
<li>Random Forest</li>
<li>K neares neighbours</li>
</ul>
```{r}
fitControl <- trainControl(method = "cv",   number = 5,
                           savePredictions = 'final',allowParallel = TRUE)

models<-caretList(beerconsumption ~., data = Train,methodList =  c("lm","rf","lasso","ridge","knn"),
             preProcess=c('center','scale'),
             trControl = fitControl)


```

```{r}
models$knn
```

```{r}
models$ridge
```

<h5> <b>Linear regression and Ridge models provided the lowest root mean squared error overall</b></h5>
```{r}
model_results <- data.frame(
 LM = mean(models$lm$results$RMSE),
 KNN = mean(models$knn$results$RMSE),
 RF = mean(models$rf$results$RMSE),
 LASSO = mean(models$lasso$results$RMSE),
 RIDGE = mean(models$ridge$results$RMSE)
 )
print(model_results)
```



